---
title: "EDA"
author: "Ivan Kostine, Niccolo Cherubini, Charlène Khairallah"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: cosmo
    highlight: rstudio
    toc: true
    toc_float: true
    code_folding: hide
---

Our data set is composed of 21 COT reports of commodities and an excel file containing their daily future prices (AJOUTER LES DATES= FROM TO). Our goal is to forecast the next COT report, a report that is done weekly. We will do so for the following commodities: Cotton, Cocoa, Coffee and Sugar. We expect to find similarities between those commodities as they belong to the same category named "softs". This means that they are agricultural products. Nevertheless, we will add the Crude Oil commodity as soft commodities should be dependent to. Indeed, Oil is the energy that farmers need for their machines and so, it can explain variations in their COT reports. 

To do our forecast, we need to understand the behavior of our commodities and the dependence that could exist between them. First, we will focus our analysis on the variables of the COT reports, then we will analyse their daily future prices. 

**1. Understanding the commodities and their relationship**
Our exploratory data analysis begin with the global understanding of each commodities through one variable that is Open Interest. We do univariate vizualization of this variable for each commodity. 

The open interest is known as a variable that contains a lot of information from the other COT variables. We will developed this later in our analysis.

*Open Interest of each commodities*
```{r Plotting open interests}
# Create and display interactive area charts for "Open Interest" for each commodity
selected_commodities <- c("Cotton", "Coffee", "Sugar", "CrudeOil", "Cocoa")

for (commodity_name in selected_commodities) {
  current_data <- commodity_data_list[[commodity_name]]
  
  open_interest_chart <- current_data %>%
    ggplot(aes(x = Date, y = `Open_Interest`)) +
    ggtitle(paste("Open interest of ", commodity_name))+
    geom_area(fill = "#69b3a2", alpha = 0.5) +
    geom_line(color = "#69b3a2") +
    ylab("Open Interest") +
    theme_ipsum()
    
  interactive_area_chart <- ggplotly(open_interest_chart)
  
  # Display the interactive area chart
  print(interactive_area_chart)
}


```
The graphs shows a similar trend between Cocoa and Coffee. Also, it shows a similar trend between CrudeOil, Sugar and Cotton. To be sure, we do a bivariate vizualization for Cocoa and Coffee, and a multivariate one for the other commodities. 

```{r}
# Assuming commodity_data_list is a list containing data frames for each commodity

# Extract data for Cocoa and Coffee
cocoa_data <- commodity_data_list$Cocoa
coffee_data <- commodity_data_list$Coffee

# Fit linear regression lines
cocoa_lm <- lm(Open_Interest ~ Date, data = cocoa_data)
coffee_lm <- lm(Open_Interest ~ Date, data = coffee_data)

# Create a scatter plot with trend lines
comparison_plot <- ggplot() +
  geom_point(data = cocoa_data, aes(x = Date, y = Open_Interest, color = "Cocoa"), alpha = 0.5) +
  geom_point(data = coffee_data, aes(x = Date, y = Open_Interest, color = "Coffee"), alpha = 0.5) +
  geom_smooth(data = cocoa_data, aes(x = Date, y = Open_Interest, color = "Cocoa"), method = "lm", se = FALSE) +
  geom_smooth(data = coffee_data, aes(x = Date, y = Open_Interest, color = "Coffee"), method = "lm", se = FALSE) +
  ggtitle("Comparison of Open Interest: Cocoa vs. Coffee") +
  ylab("Open Interest") +
  xlab("Date") +
  theme_minimal()

# Display the plot
print(comparison_plot)

```
This scatterplot seems to show a common trend between Coffee and Cocoa. 

```{r}
# Assuming commodity_data_list is a list containing data frames for each commodity

# Extract data for Cotton, Sugar, and Crude Oil
cotton_data <- commodity_data_list$Cotton
sugar_data <- commodity_data_list$Sugar
oil_data <- commodity_data_list$CrudeOil

# Fit linear regression lines
cotton_lm <- lm(Open_Interest ~ Date, data = cotton_data)
sugar_lm <- lm(Open_Interest ~ Date, data = sugar_data)
oil_lm <- lm(Open_Interest ~ Date, data = oil_data)

# Create a scatter plot with trend lines
comparison_plot <- ggplot() +
  geom_point(data = cotton_data, aes(x = Date, y = Open_Interest, color = "Cotton"), alpha = 0.5) +
  geom_point(data = sugar_data, aes(x = Date, y = Open_Interest, color = "Sugar"), alpha = 0.5) +
  geom_point(data = oil_data, aes(x = Date, y = Open_Interest, color = "Crude Oil"), alpha = 0.5) +
  geom_smooth(data = cotton_data, aes(x = Date, y = Open_Interest, color = "Cotton"), method = "lm", se = FALSE) +
  geom_smooth(data = sugar_data, aes(x = Date, y = Open_Interest, color = "Sugar"), method = "lm", se = FALSE) +
  geom_smooth(data = oil_data, aes(x = Date, y = Open_Interest, color = "Crude Oil"), method = "lm", se = FALSE) +
  ggtitle("Comparison of Open Interest: Cotton vs. Sugar vs. Crude Oil") +
  ylab("Open Interest") +
  xlab("Date") +
  theme_minimal()

# Display the plot
print(comparison_plot)

```
From this scatterplot, Crude Oil, Cotton and Sugar seems to follow the same trend but in different proportions. 

(AJOUTER DE LANALYSE)

**2. Understanding the variables in the COT report**
As mentioned before, the open interest variable is very important to understand the variation in the COT reports. From the following dygraphs, we can see that Open interest, total reportable shorts and total reportable longs are highly correlated.

```{r Plotting all variables for each commodity}
# Create and display interactive area charts for "Open Interest" for each commodity
selected_commodities <- c("Cotton", "Coffee", "Sugar", "CrudeOil", "Cocoa")

for (commodity_name in selected_commodities) {
  current_data <- commodity_data_list[[commodity_name]]
  
  # Create xts object from your data
  PLS <- xts(current_data, order.by = current_data$Date)
  
  # Create dygraph
  P <- dygraph(PLS, main = paste(commodity_name))
  
  # Display the dygraph
  print(P)
}


```

Using this visualization, we have a better understanding of each variable time series.  
```{r Plotting time series for each variable of each commodity}
# Create and display interactive area charts for "Open Interest" for each commodity
selected_commodities <- c("Cotton", "Coffee", "Sugar", "CrudeOil", "Cocoa")

for (commodity_name in selected_commodities) {
  current_data <- commodity_data_list[[commodity_name]]
  
  COT_data_long <- gather(current_data, key = "Variable", value = "Value", -Date)
  
  plotTimeseries <- ggplot(data = COT_data_long, aes(x = Date, y = Value)) +
    geom_line() +
    facet_wrap(~Variable, scales = "free_y") +
    labs(
      title = paste("Time Series Plots for COT", commodity_name),  # Include commodity name in the title
      x = "Date",
      y = "Value"
    )
  
  print(plotTimeseries)
}


```
Besides the common trends of Open interest, Total Reportable Longs and Total Reportable Shorts,
we notice that Money Manager Spread is sharing some similarities with these variables too.

With the following graphs, we aimed to see the relationship between producer Longs and Shorts. Most commodities seem to display positive correlation between Producer Longs and Producer Shorts. But Cocoa and Coffee share similarities as the two variables show signs of negative correlations. 
```{r Plotting longs vs. shorts}

selected_commodities <- c("Cotton", "Coffee", "Sugar", "CrudeOil", "Cocoa")

for (commodity_name in selected_commodities) {
  current_data <- commodity_data_list[[commodity_name]]
  
  # Create xts object with selected variables
  PLS <- xts(data.frame(
    Date = current_data$Date,
    producer_short = current_data$`Producer_Longs`,
    producer_long = current_data$`Producer_Shorts`
  ), order.by = current_data$Date)
  
  # Create dygraph
  P <- dygraph(PLS, main = paste("Producer Longs vs. Shorts -", commodity_name))
  
  # Display the dygraph
  print(P)
}

```

With correlation heatmaps, we clearly see the correlation seen above, between the variables of the COT reports. We notie that the Cocoa do not exhibit any negative correlation and Coffee do exhibit very few ones compared to Sugar for example.
```{r Plotting correlation heatmaps}

selected_commodities <- c("Cotton", "Coffee", "Sugar", "CrudeOil", "Cocoa")

for (commodity_name in selected_commodities) {
  current_data <- commodity_data_list[[commodity_name]] 
  
  cormat = round(cor(current_data[,-1]),2)
    melted_cormat <- melt(cormat) %>% rename("Variable_1"="Var1", "Variable_2"="Var2")
    plot = ggplot(data = melted_cormat, aes(x=Variable_1, y=Variable_2, fill=value)) +
    ggtitle(paste("Correlation heatmap", commodity_name))+
    geom_tile(color = "white") +
    scale_fill_gradient2(high = "darkblue", low = "darkred", mid = "white", midpoint = 0, limit = c(-1,1), space = "Lab", name="Pearson\nCorrelation") +
    theme_minimal()+ 
    theme(axis.text.y = element_text(size=7))+
    theme(axis.text.x = element_text(angle = 40, vjust = 1, size = 7, hjust = 1))+
    coord_fixed()
    print(plot)
  }


```

We change the Date variable to consecutive integers because of its irregularity. This modification will help us to manipulate and analyse time series.
```{r Transforming data}

# Create a new list to store data frames with modified dates
commodity_data_list_modified <- list()

# Iterate through the original list and create a modified copy
for (commodity in commodities) {
  # Make a copy of the data from the original list
  data_copy <- data.frame(commodity_data_list[[commodity]])
  
  # Modify the Date column to consecutive integers
  data_copy$Date <- 1:nrow(data_copy)
  
  # Store the modified data in the new list
  commodity_data_list_modified[[commodity]] <- as_tsibble(data_copy, index = Date, regular = FALSE)
}


```

Then we standardize our data.

```{r Standardizing data}
# Create a new list to store standardized data frames
standardized_commodity_data_list <- list()

# Iterate through the modified data list and standardize each data frame
for (commodity in commodities) {
  # Get the modified data from the modified list
  modified_data <- commodity_data_list_modified[[commodity]]
  
  # Standardize the data excluding the "Date" column
  standardized_data <- scale(modified_data[, -1])
  
  # Create a new data frame with standardized values
  standardized_data_frame <- as.data.frame(standardized_data)
  
  # Add back the "Date" column
  standardized_data_frame$Date <- modified_data$Date
  
  # Convert to tsibble
  tsibble_data <- as_tsibble(standardized_data_frame, index = Date, regular = FALSE)
  
  # Store the standardized data in the new list
  standardized_commodity_data_list[[commodity]] <- tsibble_data
}

```

Covariance matrix:

```{r Covariance Matrix}
# Create a new list to store covariance matrices
covariance_matrices_list <- list()

# Iterate through the standardized data list and calculate the covariance matrix
for (commodity in commodities) {
  # Get the standardized data from the standardized list
  standardized_data <- standardized_commodity_data_list[[commodity]]
  
  # Exclude the "Date" column
  trading_vars <- as.matrix(standardized_data[, -17])
  
  # Calculate the covariance matrix
  cov_matrix <- cov(trading_vars)
  
  # Store the covariance matrix in the new list
  covariance_matrices_list[[commodity]] <- cov_matrix
}

covariance_matrices_list
```

PCA 


```{r PCA}
# Create a new list to store PCA results
pca_results_list <- list()

# Iterate through the covariance matrices and perform PCA
for (commodity in commodities) {
  # Get the covariance matrix from the list
  cov_matrix <- covariance_matrices_list[[commodity]]
  
  # Perform PCA using prcomp
  pca_result <- prcomp(cov_matrix, scale. = TRUE)
  
  # Store the PCA result in the new list
  pca_results_list[[commodity]] <- pca_result
}

pca_results_list

```

SCREE PLOTS

```{r Scree Plots}
# Create a new list to store scree plots
scree_plots_list <- list()

# Iterate through the PCA results and create scree plots
for (commodity in commodities) {
  pca_result <- pca_results_list[[commodity]]
  
  # Extract eigenvalues
  eigenvalues <- pca_result$sdev^2
  
  # Create scree plot
  scree_plot <- plot(1:length(eigenvalues), eigenvalues, type = "b", 
                     main = paste("Scree Plot -", commodity),
                     xlab = "Principal Component", ylab = "Eigenvalue")
  
  # Store the scree plot in the list
  scree_plots_list[[commodity]] <- scree_plot
}

#scree_plots_list
```

BIPLOTS
```{r Biplots}
# Create a new list to store biplots
biplots_list <- list()

# Iterate through the PCA results and create biplots
for (commodity in commodities) {
  pca_result <- pca_results_list[[commodity]]
  
  # Create biplot
  biplot <- biplot(pca_result, scale = 0)
  
  # Add title to the biplot
  title <- paste("Biplot -", commodity)
  title(main = title)
  
  # Store the biplot in the list
  biplots_list[[commodity]] <- biplot
}

#biplots_list
```



We want to retain the principal components associated with eigenvalues above the "elbow" in the scree plot.

VARIABLES SELECTION = ASK Professor BOLDI FOR THE METHOD PLS

```{r Variable selection}
# Create a new list to store the most significant variables
most_significant_variables_list <- list()

# Iterate through the PCA results
for (commodity in commodities) {
  pca_result <- pca_results_list[[commodity]]
  
  # Extract loadings from the PCA result
  loadings <- pca_result$rotation
  
  # Identify the most significant variable for each principal component
  most_significant_variables <- apply(loadings, 2, function(pc_loadings) {
    variable_index <- which.max(abs(pc_loadings))
    variable_name <- colnames(loadings)[variable_index]
    return(variable_name)
  })
  
  # Store the most significant variables in the list
  most_significant_variables_list[[commodity]] <- most_significant_variables
}

# Print or use the most significant variables list as needed
print(most_significant_variables_list)

```
The labels "PC14," "PC5," etc., are likely the names or labels of the original variables in your dataset. These labels indicate which variables contribute the most to the variability captured by each principal component.

NUMBER OF PCA TO SELECT


```{r PCA selection}
# Create a list to store the cumulative variance explained for each commodity
cumulative_variance_explained_list <- list()

# Iterate through the PCA results
for (commodity in commodities) {
  pca_result <- pca_results_list[[commodity]]
  
  # Access eigenvalues and calculate cumulative variance
  eigenvalues <- pca_result$sdev^2
  cumulative_variance <- cumsum(eigenvalues) / sum(eigenvalues)
  
  # Store the cumulative variance explained in the list
  cumulative_variance_explained_list[[commodity]] <- cumulative_variance
}

# Select the number of principal components based on cumulative variance explained
# Adjust the threshold as needed
threshold <- 0.95
selected_components_list <- lapply(cumulative_variance_explained_list, function(cum_var) {
  which(cum_var >= threshold)[1]
})

# Print or use the selected components list as needed
print(selected_components_list)

```

```{r Forcasting using ETS}
library(tsibble)
library(fable)
library(ggplot2)

# Assuming COT_Cotton is your tsibble
# Assuming you have a column 'Open_Interest' and 'Date'

# Convert 'Date' to Date format
COT_Cotton$Date <- as.Date(COT_Cotton$Date)
newdata <- COT_Cotton %>%
  select(Date, `Open_Interest`, `Total_Reportable_Longs` , `Total_Reportable_Shorts` )


# Create a tsibble with regular dates starting from January 2006
regular_tsibble <- tsibble(date = seq(from = as.Date("2006-01-01"), by = "1 week", length.out = nrow(newdata)))

# Merge COT_Cotton and regular_tsibble based on nothing
merged_data <- bind_cols(newdata, regular_tsibble)

merged_data <- as_tsibble(merged_data%>% select(-Date), key = NULL)

# Fit the ETS model using the merged data
fit <- merged_data %>%
  model(ETS(`Open_Interest` ~ error("A") + trend("N") + season("N")))

# Forecast for the next observations
fc <- fit %>%
  forecast(h = 1)
fc
 
# Plot the forecast
fc |>
  autoplot(merged_data) +
   geom_line(aes(y = .fitted), col = "#D55E00", data= augment(fit)) +
   labs(y = "Open Interest", title = "Forecast: Cotton Open Interest") +
   guides(colour = "none")


```
Multivariate F avec Coffee, cocoa et sugar ? ESSAI FAIL(pb data regroupement)
peut etre multivariate avec une commo et plusieurs variables de son COT (A FAIRE)
Also, test ARIMA à faire.
```{r}
# library(tsibble)
# library(fable)
# library(dplyr)

# # Assuming commodity_data_list is your list of datasets
# 
# # Extract Open_Interest for Cocoa, Sugar, and Coffee
# open_interest_data <- commodity_data_list %>%
#   bind_rows(.id = "commodity") %>%
#   filter(commodity %in% c("Cocoa", "Sugar", "Coffee")) %>%
#   select(-Date) %>%
#   rename(Open_Interest = `Open Interest`)
# 
# # Create a regular tsibble with dates starting from January 2006
# regular_tsibble <- tsibble(date = seq(from = as.Date("2006-01-01"), by = "1 week", length.out = nrow(open_interest_data)))
# 
# # Bind the regular tsibble with the Open Interest data
# merged_data2 <- bind_cols(regular_tsibble, open_interest_data)
# 
# # Fit a multivariate ETS model
# fit2 <- merged_data2 %>%
#   model(ETS(Open_Interest ~ error("A") + trend("N") + season("N") | commodity))
# 
# # Forecast for the next observations
# fc2 <- fit2 %>%
#   forecast(h = 1)
# 
# # Plot the forecast
# fc2 |>
#   autoplot(merged_data2) +
#    geom_line(aes(y = .fitted), col = "#D55E00", data= augment(fit2)) +
#    labs(y = "Open Interest", title = "Forecast: Multivariate ETS for Open Interest") +
#    guides(colour = "none")
# 


```



ETS des variables correlees du Cotton
```{r ETS with cotton}


COT_Cotton$Date <- as.Date(COT_Cotton$Date)

newdata2 <- COT_Cotton %>% select(Date, `Open_Interest`, `Total_Reportable_Longs`,`Total_Reportable_Shorts`)

# Create a tsibble with regular dates starting from January 2006
regular_tsibble3 <- tsibble(date = seq(from = as.Date("2006-01-01"), by = "1 week", length.out = nrow(newdata2)))

# Merge COT_Cotton and regular_tsibble based on nothing
merged_data3 <- bind_cols(newdata2, regular_tsibble3)

merged_data3 <- as_tsibble(merged_data3 %>% select(-Date), key = NULL)


# Fit the ETS model using the merged data
fit3 <- merged_data3 %>%
  model(
    ARIMA(Open_Interest ~ Total_Reportable_Longs + Total_Reportable_Shorts))
# 
# # Forecast for the next observations
# fc3 <- fit3 %>%
#   forecast(h = 1)
# 
# fc3
#  
# # Plot the forecast
# fc3 |>
#   autoplot(merged_data3) +
#    geom_line(aes(y = .fitted), col = "#D55E00", data= augment(fit3)) +
#    labs(y = "Open Interest", title = "Forecast: Cotton Open Interest based on multivariate ETS") +
#    guides(colour = "none")
```

